---
title: "R Notebook"
output: html_notebook
---

Maybe try Poisson?

Need the original dataset... not 0's and 1's
```{r}
####R script exploration####

#import the data
Groundfish <- read.csv("ai2014_2018.csv")
names(Groundfish)

#load some packages
library(ggplot2)
library(dplyr)
library(tidyverse)
library(dplyr)

#filter some data
unique(Groundfish$COMMON)
Sablefish <- Groundfish %>% filter(COMMON == "sablefish")
names(Sablefish)
#I'm interested in NUMCPUE as my covariate

#let's load that geostatistical package we used for scallops
##and use that to make a geodataframe
##and plot just like we did for scallops, with the covariate as sablefish number
??as.geodata
library(geoR)
##seems to have loaded, got some weird warnings tho
data("Ksat") #just looking at this data structure


#the challenge of creating a geodata object so I can plot it
length(Sablefish$LATITUDE)
names(Sablefish)
Sablefish_relevant <- Sablefish[c(2,1,8)] #got the relevant data frame here
###you may want to add more
#Sable.matrix <- as.matrix(data=c(Sablefish$LONGITUDE, Sablefish$LATITUDE, ),nrow=162)
#make a matric
Sable_matrix <- as.matrix(Sablefish_relevant) #that worked!!
Sable_geo <- as.geodata(Sable_matrix) #now create the geodata object

##WARNING!! I pulled out all examples of sablefish. There are surveys with 0 sablefish that I did not
##acknoledge at this point. Those 0's are probably important


##which surveys do we have?
names(Sablefish)
Sablefish_survey_df <- Sablefish[c(16,17)]
Sablefish
#surveys total
Groundfish$HAUL


#unite
SableID1 <- unite(Sablefish_survey_df, ID_new, CRUISE:HAUL)
names(Sablefish_survey_df)
head(Sablefish_survey_df)

Groundfish_survey_ID <- Groundfish[c(16,17)]

GroundID1<- unite(data=Groundfish_survey_ID, col=ID_new, CRUISE:HAUL)

GroundID <-as.factor(GroundID1$ID_new)
SableID <- as.factor(SableID1$ID_new)
nosable <- unique(setdiff(GroundID, SableID))
length(unique(GroundID))
length(nosable)
length(unique(SableID)) #not the same as sable ID. I do not know why


#seemed to work. not let's find the coords for nosable
Groundfish$ID_new <- GroundID #a factor
names(Groundfish)
head(Groundfish)
#does intersect work here
#nosable.df <- Groundfish %>% intersect(GroundID, nosable)
nosable.df <- nosable %in% Groundfish$ID_new #HELP

Groundfish_locations_nosable <- Groundfish %>% filter(ID_new %in% c(nosable))
#THANK YOU MATT!!
names(Groundfish_locations_nosable)
Groundfish_locations_nosable_coords <- unique(Groundfish_locations_nosable[c(2,1,18)])
Groundfish_locations_nosable_coords2 <- unique(Groundfish_locations_nosable[c(2,1)])
length(Groundfish$LATITUDE)
length(Groundfish_locations_nosable_coords$LATITUDE)
length(unique(Groundfish_locations_nosable_coords$ID_new))
length(Groundfish_locations_nosable_coords2$LATITUDE)

head(Groundfish_locations_nosable_coords)
names(Groundfish_locations_nosable_coords)
head(Sable_matrix)
Groundfish_locations_nosable_coords$NUMCPUE <- rep(0, 993)
Nosable_matrix_prep <- Groundfish_locations_nosable_coords[-3]

#so theres 162 sablefish hauls and
##993 hauls without sablefish if we use the Groundfish_locations_nosable_coords dataset
##let's combine the dataframe so that we have sablefish coords and our zeros!!

head(Sable_matrix)
Sable_0_matrix <- as.matrix(Nosable_matrix_prep)
#Sable_matrix[,1] <- abs(Sable_matrix[,1]) #this was not the right step
#Sable_0_matrix[,1] <- abs(Sable_0_matrix[,1] ) #tis was not the right step

Sable_matrix_combined <- rbind(Sable_matrix, Sable_0_matrix)

Sable_geo_P <- as.geodata(Sable_matrix_combined)
Sable_geo_P
plot(Sable_geo_P)

lons <- Sable_geo_P$coords[,1]
lons <- ifelse(lons<0, lons+180, lons-180)
Sable_geo_P$coords[,1] <- lons
plot(Sable_geo_P) #the poisson geodataframe is set up
```

Step1: jitter coords fir Sable_geo_P
```{r}
names(Sable_geo_P)
coord.df.P <-data.frame(Sable_geo_P$coords[,1], Sable_geo_P$coords[,2])
Sable_coord_jitter.P <- jitterDupCoords(coord.df.P, max=0.01)

Sable_coord_jitter.P$data <- Sable_geo_P$data

Sable_coord_P_jitter_gdf <- as.geodata(Sable_coord_jitter.P)
dup.coords(Sable_coord_P_jitter_gdf)
```

Step2: Fit an empirical semivariogram and a WLS semivriogram to get param ests.

```{r}
library(dplyr)
Sable_coord_P_jitter_gdf_log <- Sable_coord_P_jitter_gdf
Sable_coord_P_jitter_gdf_log$data <- log(Sable_coord_P_jitter_gdf$data + 0.01)

boxwid <- 1

my_grid <- pred_grid(  
  c(-10, 15),
  c(50, 55),
  by=boxwid
) #These are the same as the grid before



dim(my_grid)

my_variog_sable_jitter_P<- variog(Sable_coord_P_jitter_gdf_log, trend="1st") 
#my_variog_sable_2<- variog(Sable_geo_binary, trend="2nd") 

plot(my_variog_sable_jitter_P)
#plot(my_variog_sable_2)
#
#both look weird
##gonna roll with trend=1st for now.

##Variofit for the WLS estimate
my_WLS_sable_jitter_P <- variofit(my_variog_sable_jitter_P,
                        ini.cov.pars = c(3, 5), #sill, then range estimates
                        cov.model="exponential",
                        fix.nugget=F,
                        max.dist = 300 #plot gets weird after 300
                        )

##plot it over your variog
plot(my_variog_sable_jitter_P)
lines(my_WLS_sable_jitter_P) #looks like a good fit to me.

names(my_WLS_sable_jitter_P)
sable.cov.pars_P <- my_WLS_sable_jitter_P$cov.pars #sigsq, then phi
sable_WLS_sigsq_P <- sable.cov.pars_P[1]
sable_WLS_phi_P <- sable.cov.pars_P[2]
sable_WLS_tao_sq_P <- my_WLS_sable_jitter_P$nugget
sable_WLS_w_P <- sable_WLS_tao_sq_P/sable_WLS_sigsq_P
```

Step 2.5 ML and REML parameter estimates (I'm skipping this for now)

Step 3: Poisson regression, for beta values
```{r}
#using gdf: Sable_coord_P_jitter_gdf_log
#df wrangle
names(Sable_coord_P_jitter_gdf_log)
names(Sable_coord_P_jitter_gdf)
#df <- data.frame(Sable_coord_P_jitter_gdf_log)
#df$LONGITUDE <- Sable_coord_P_jitter_gdf_log$Sable_geo_P.coords...1.
#df$LATUTUDE <- Sable_coord_P_jitter_gdf_log$Sable_geo_P.coords...2.
#names(df)
  
  
#use the logged or unlogged datas here?
#try the unlogged data
fit_P <- glm(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2. , family=poisson(link="log"), data=Sable_coord_P_jitter_gdf)
beta.starting_P <- coef(fit_P)
beta.tuning_P <- t(chol(vcov(fit_P)))

summary(fit_P)

#try the logged data?
####


#got my betas
```


Step 4: run the spGLM
m.jitter.P (saved as p2.rds) is the winning model right now
```{r}


m.jitter.P <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=Sable_coord_P_jitter_gdf$coords, data=Sable_coord_P_jitter_gdf, n.samples = 5000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P,"sigma.sq"=sable_WLS_sigsq_P, "beta"=beta.starting_P, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_P , "w" = 0.05)) 


saveRDS(m.jitter.P, "m_jitter_P2.rds")
names(m.jitter.P)
plot(mcmc(m.jitter.P$p.beta.theta.samples))

m.jitter.P$acceptance
names(m.jitter.P)

```

This will be my last attempt:BAD
Report interval Metrop. Acceptance rate: 0.00%: Does this mean flat lining?
BAD  BAD  Does our one outlier matter? remove outlier from dataset?
```{r}
m.jitter.P_3 <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=Sable_coord_P_jitter_gdf$coords, data=Sable_coord_P_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P,"sigma.sq"=sable_WLS_sigsq_P, "beta"=beta.starting_P, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_P , "w" = 0.025)) 


#saveRDS(m.jitter.P_3, "m_jitter_P3.rds")
#names(m.jitter.P)
plot(mcmc(m.jitter.P_3$p.beta.theta.samples))
```
```{r}
m.jitter.P_4 <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=Sable_coord_P_jitter_gdf$coords, data=Sable_coord_P_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P,"sigma.sq"=sable_WLS_sigsq_P, "beta"=beta.starting_P, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.5, "sigma.sq"=0.5, "beta"= beta.tuning_P , "w" = 0.5)) 
```

```{r}
hist(log(Sable_coord_P_jitter_gdf$data+1))
```
Jitter 5. Log + 1. Fixing the logs...


```{r}

```

```{r}
P_new <- Sable_coord_P_jitter_gdf
P_new$data <- Sable_coord_P_jitter_gdf$data+1
#########

fit_P2 <- glm(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2. , family=poisson(link="log"), data=P_new)
beta.starting_P_new <- coef(fit_P2)
beta.tuning_P_new <- t(chol(vcov(fit_P2)))

summary(fit_P2)

##do I need to change the variogram too? to +1 for the data? I'm not thinking that I do....
###CHANGING STEP 2
Sable_coord_P_jitter_gdf_log_new <- Sable_coord_P_jitter_gdf
Sable_coord_P_jitter_gdf_log_new$data <- log(Sable_coord_P_jitter_gdf$data + 1)

my_variog_sable_jitter_P_new<- variog(Sable_coord_P_jitter_gdf_log_new, trend="1st") 
#my_variog_sable_2<- variog(Sable_geo_binary, trend="2nd") 

plot(my_variog_sable_jitter_P_new)
#plot(my_variog_sable_2)

##Variofit for the WLS estimate
my_WLS_sable_jitter_P_new <- variofit(my_variog_sable_jitter_P_new,
                        ini.cov.pars = c(0.2, 7), #sill, then range estimates
                        cov.model="exponential",
                        fix.nugget=F,
                        max.dist = 300 #plot gets weird after 300
                        )

##plot it over your variog
plot(my_variog_sable_jitter_P_new)
lines(my_WLS_sable_jitter_P_new) #looks like a good fit to me.

names(my_WLS_sable_jitter_P_new)
sable.cov.pars_P_new <- my_WLS_sable_jitter_P_new$cov.pars #sigsq, then phi
sable_WLS_sigsq_P_new <- sable.cov.pars_P_new[1]
sable_WLS_phi_P_new <- sable.cov.pars_P_new[2]
sable_WLS_tao_sq_P_new <- my_WLS_sable_jitter_P_new$nugget
sable_WLS_w__newP <- sable_WLS_tao_sq_P_new/sable_WLS_sigsq_P_new
#########
######

##LAST 2 ATTEMPTS!!

m.jitter.P_5 <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=P_new$coords, data=P_new, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P_new,"sigma.sq"=sable_WLS_sigsq_P_new, "beta"=beta.starting_P_new, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.5, "sigma.sq"=0.5, "beta"= beta.tuning_P_new , "w" = 0.5))
plot(mcmc(m.jitter.P_5$p.beta.theta.samples))
#maybe change to smaller in V6

m.jitter.P_6 <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=P_new$coords, data=P_new, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P_new,"sigma.sq"=sable_WLS_sigsq_P_new, "beta"=beta.starting_P_new, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.25, "sigma.sq"=0.25, "beta"= beta.tuning_P_new , "w" = 0.25)) 

plot(mcmc(m.jitter.P_6$p.beta.theta.samples))
#next try smaller: 0.11 instead of 0.25 for tuning. maybe keep w at 0.5

##might have set this up wrong... don't use log-tranformed data alreadyy??
m.jitter.P_7 <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=P_new$coords, data=P_new, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P_new,"sigma.sq"=sable_WLS_sigsq_P_new, "beta"=beta.starting_P_new, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.123, "sigma.sq"=0.123, "beta"= beta.tuning_P_new , "w" = 0.5)) 

plot(mcmc(m.jitter.P_7$p.beta.theta.samples))


#maybe try log tranforming the data FIRST, that way we have zeros. I think this is the way
#careful to sue th right P_new, the already logged P_new (not right in the order right now)

P_new$data <- log(Sable_coord_P_jitter_gdf$data+1)

m.jitter.P_8 <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=P_new$coords, data=P_new, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P_new,"sigma.sq"=sable_WLS_sigsq_P_new, "beta"=beta.starting_P_new, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_P_new , "w" = 0.2)) 

plot(mcmc(m.jitter.P_8$p.beta.theta.samples))
##this looks promising!!!!!!!!!!!
#saveRDS(m.jitter.P_8, "m_jitter_P8.rds")

##widening acceptanes again..?
m.jitter.P_9 <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=P_new$coords, data=P_new, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P_new,"sigma.sq"=sable_WLS_sigsq_P_new, "beta"=beta.starting_P_new, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.025, "sigma.sq"=0.025, "beta"= beta.tuning_P_new , "w" = 0.2)) #change these?? to 0.05 for w?

plot(mcmc(m.jitter.P_9$p.beta.theta.samples)) ##BAD...


m.jitter.P_10 <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=P_new$coords, data=P_new, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=2,"sigma.sq"=sable_WLS_sigsq_P_new, "beta"=beta.starting_P_new, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 7), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_P_new , "w" = 0.2)) 

plot(mcmc(m.jitter.P_10$p.beta.theta.samples))
```

```{r}
m.jitter.P_t <- spGLM(data~Sable_geo_P.coords...1. + Sable_geo_P.coords...2. +Sable_geo_P.coords...1.:Sable_geo_P.coords...2., family="poisson", coords=P_new$coords, data=P_new, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_P_new,"sigma.sq"=sable_WLS_sigsq_P_new, "beta"=beta.starting_P_new, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01,  3), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_P_new , "w" = 0.2)) 

plot(mcmc(m.jitter.P_t$p.beta.theta.samples))
```

