---
title: "R Notebook"
output: html_notebook
---

Bulding off of R_nb_sable_2 ( and Stats_proj_sablefish.R)
04/28/22


Load in models that ran
```{r}
m1<- readRDS("jittered_mod.rds") # It ran but I didnt know what I was doing.
m2 <- readRDS("jittered_mod_2.rds") #1000 iterations
m3 <-readRDS("m_jitter_3.rds") #5000 iterations

m.jitter.3<- m3
```

plot the models that ran: m2

```{r}
m.jitter.2 <- m2

p.beta.theta.samples <- m.jitter.2$p.beta.theta.samples
plot(p.beta.theta.samples)
plot(mcmc(cbind(p.beta.theta.samples))) #plotting it, looks terrible. Why? This plot works, tho!

int <-p.beta.theta.samples[,1] #the intercept
#plot(mcmc(int)) #plotting a terrible, not grassy plot. Why?


#smaller: 
my_which_2 <- seq(from=100, to=1000, by=10)
my_theta_draws_2 <- p.beta.theta.samples[my_which_2]
str(my_theta_draws_2)
summary(my_theta_draws_2)

names(int)
my_int <- int[my_which_2]
plot(int)
#int$statistics

plot(mcmc(cbind(int)))


#ok, we figured out intercept, now plot the other traceplots
beta1 <- p.beta.theta.samples[,2]
beta2 <- p.beta.theta.samples[,3]
beta3 <- p.beta.theta.samples[,4]

my_beta0 <- my_int
my_beta1 <- beta1[my_which_2]
my_beta2 <- beta2[my_which_2]
my_beta3 <- beta3[my_which_2]

plot(mcmc(cbind(my_beta0, my_beta1, my_beta2, my_beta3)), smooth=F, density=F) #floor on B0, ceiling on B1, ceiling on B2, ceiling on B3
plot(my_beta2)

```


plot the models that ran: m3
```{r}
p.beta.theta.samples.3 <- m.jitter.3$p.beta.theta.samples

names(m.jitter.3)
m.jitter.2$acceptance # dont understand why this one is higher....
m.jitter.3$acceptance # small acceptance

beta0_3 <- p.beta.theta.samples.3[,1]
beta1_3 <- p.beta.theta.samples.3[,2]
beta2_3 <- p.beta.theta.samples.3[,3]
beta3_3 <- p.beta.theta.samples.3[,4]

my_which_3 <- seq(from=100, to=5000, by=100)

my_beta0_3 <- beta0_3[my_which_3]
my_beta1_3 <- beta1_3[my_which_3]
my_beta2_3 <- beta2_3[my_which_3]
my_beta3_3 <- beta3_3[my_which_3]

plot(mcmc(cbind(my_beta0_3, my_beta1_3, my_beta2_3, my_beta3_3)), smooth=F, density=F) 
plot(p.beta.theta.samples.3, smooth=F)
```



How to improve the next model version?
Need a wider metropolis proposal. How to do that here? Tuning?
I'll try setting larger tuning values for this one?
Maybe increase the variance? It's 10 right now. Consider increasing to 1000.
```{r}
library(geoR)
library(ggplot2)
library(coda)
library(spBayes)
m.jitter.4 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #inc iterations from 1000
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(1, 2), "sigma.sq.IG"=c(2,1),                                                                                                         "beta.Normal"=list(c(0,0,0,0),c(10,10,10,10))), #gonna want to make these 1000...
             tuning=list("phi"=0.2, "sigma.sq"=0.2, "beta"= beta.tuning_j_3, "w" = 0.2)) #made tuning lower to see waht happens

##have the output tell me: what are the accep
m.jitter.4$acceptance #2.3 way too small
m.jitter.3$acceptance

```

m.jitter.5
Lowering the tuning even more!!
And raise the variance of the priors
(when I use family=binomial, am I assuming a normal distribution (stupid question...))
```{r}
m.jitter.5 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(1, 2), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.005, "sigma.sq"=0.005, "beta"= beta.tuning_j_3, "w" = 0.005)) #made tuning EVEN lower to see what happens
saveRDS(m.jitter.5, "m_jitter_5.rds")

#these dont have to be the same. We can make one of them smaller
```

Acceptance is still not high enough
```{r}
m.jitter.5$acceptance
```

Even smaller tuning! If this doesnt work, change beta.tuning_j_3
...."nu" is an option. What is nu?
```{r}
m.jitter.6 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(1, 2), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.005, "sigma.sq"=0.005, "beta"= beta.tuning_j_3, "w" = 0.005)) #made tuning EVEN lower to see what happens
saveRDS(m.jitter.6, "m_jitter_6.rds")
```


Ok so m.jitter.5 looks better than m.jitter.6. But I don't know why.
```{r}
m.jitter.5$acceptance
m.jitter.6$acceptance

print(summary(window(m.jitter.5$p.beta.theta.samples)))

plot(mcmc(m.jitter.5$p.beta.theta.samples))


plot(mcmc(m.jitter.6$p.beta.theta.samples))

library(ggplot2)
class(Sable_coord_binary_jitter_gdf)
Test.df <- data.frame(Sable_coord_binary_jitter_gdf)
names(Test.df)
ggplot(Test.df) + aes(x=Sable_geo_df_binary.LONGITUDE, y=Sable_geo_df_binary.LATITUDE) + geom_point()

plot(m.jitter.5$p.beta.theta.samples[,2], m.jitter.5$p.beta.theta.samples[,3])
scatterplot(m.jitter.5$p.beta.theta.samples[,2], m.jitter.5$p.beta.theta.samples[,3])
ggplot() + aes(m.jitter.5$p.beta.theta.samples[,2], m.jitter.5$p.beta.theta.samples[,3]) + geom_point()
length(unique(m.jitter.5$p.beta.theta.samples[,2]))
```

In this iteration:
Change sigsq and phi
and their tuning
changed phi uniform start and variance. Maybe reduce pri.unif variance a bit
```{r}
m.jitter.7 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=1.3,"sigma.sq"=300, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 2), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.0025, "sigma.sq"=0.0025, "beta"= beta.tuning_j_3, "w" = 0.005)) #
saveRDS(m.jitter.7, "m_jitter_7.rds")
```



```{r}
plot(mcmc(m.jitter.7$p.beta.theta.samples))
```
```{r}
m.jitter.8 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=0.9,"sigma.sq"=50, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 2), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.001, "sigma.sq"=0.001, "beta"= beta.tuning_j_3, "w" = 0.0025)) #
saveRDS(m.jitter.8, "m_jitter_8.rds")
```

```{r}
plot(mcmc(m.jitter.8$p.beta.theta.samples))
```
```{r}
m.jitter.9 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=0.9,"sigma.sq"=15, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 1), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.0005, "sigma.sq"=0.0005, "beta"= beta.tuning_j_3, "w" = 0.001)) #
saveRDS(m.jitter.9, "m_jitter_9.rds")
```



```{r}
plot(mcmc(m.jitter.9$p.beta.theta.samples))
```

Alternative phi:0.9
```{r}


m.jitter.10 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.48, 2.48), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_j_3, "w" = 0.05)) 

#
saveRDS(m.jitter.10, "m_jitter_10.rds")

```


```{r}
plot(mcmc(m.jitter.10$p.beta.theta.samples))
#ok....soemthing worked. Seemed to respond to the changing phi uniform dist. Try doing this same thing but...longer??
```
vertion 10 but 5000 times!
```{r}
m.jitter.11 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 5000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=2.5,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(1, 4), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_j_3, "w" = 0.05)) 

#
saveRDS(m.jitter.11, "m_jitter_11.rds")
```


```{r}
plot(mcmc(m.jitter.11$p.beta.theta.samples))
```


```{r}
m.jitter.12 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 7), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_j_3, "w" = 0.05)) 

#
#saveRDS(m.jitter.12, "m_jitter_12.rds")
```

```{r}
plot(mcmc(m.jitter.12$p.beta.theta.samples))
```

```{r}
m.jitter.13 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=3,"sigma.sq"=60, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 7), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_j_3, "w" = 0.05)) 

#
saveRDS(m.jitter.13, "m_jitter_13.rds")
```

```{r}
plot(mcmc(m.jitter.13$p.beta.theta.samples))
```

Refresh to see whats up
```{r}
names(Sable_coord_binary_jitter_gdf)
#names()
m.jitter.13 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 2.5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.25, "sigma.sq"=0.25, "beta"= beta.tuning_j_3, "w" = 0.25)) 

#
saveRDS(m.jitter.13, "m_jitter_13.rds")
```

```{r}
plot(mcmc(m.jitter.13$p.beta.theta.samples))
```

Changed mcmc to smaller
```{r}
names(Sable_coord_binary_jitter_gdf)
#names()
m.jitter.14 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 2.5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.25, "sigma.sq"=0.25, "beta"= beta.tuning_j_3, "w" = 0.25)) 

#
saveRDS(m.jitter.14, "m_jitter_14.rds")
```


```{r}
plot(mcmc(m.jitter.14$p.beta.theta.samples))
```
raise phi, lower mcmc 
```{r}
#names(Sable_coord_binary_jitter_gdf)
#names()
m.jitter.15 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.1, "sigma.sq"=0.1, "beta"= beta.tuning_j_3, "w" = 0.1)) 

#
saveRDS(m.jitter.15, "m_jitter_15.rds")
```


```{r}
plot(mcmc(m.jitter.15$p.beta.theta.samples))
```
```{r}
m.jitter.15 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.1, "sigma.sq"=0.1, "beta"= beta.tuning_j_3, "w" = 0.1)) 

#
saveRDS(m.jitter.15, "m_jitter_15.rds")
```

Make one with altered beta
```{r}
m.jitter.16 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.1, "sigma.sq"=0.1, "beta"= c(0.1, 0.1, 0.1, 0.1) , "w" = 0.1)) 

#
saveRDS(m.jitter.16, "m_jitter_16.rds")
```

```{r}
plot(mcmc(m.jitter.16$p.beta.theta.samples))
```
```{r}
m.jitter.17 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.1, "sigma.sq"=0.1, "beta"= c(0.1, 0.1, 0.01, 0.1) , "w" = 0.1)) 

#
saveRDS(m.jitter.17, "m_jitter_17.rds")
```

```{r}
plot(mcmc(m.jitter.17$p.beta.theta.samples))
```

```{r}
m.jitter.18 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=4,"sigma.sq"=100, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 7), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= c(0.05, 0.05, 0.005, 0.05) , "w" = 0.05)) 

#
saveRDS(m.jitter.18, "m_jitter_18.rds")
```

```{r}
plot(mcmc(m.jitter.18$p.beta.theta.samples))
```
```{r}
m.jitter.19 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=4,"sigma.sq"=100, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.03, 7), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= c(0.05, 0.05, 0.005, 0.05) , "w" = 0.05)) 

#
saveRDS(m.jitter.19, "m_jitter_19.rds")
```

```{r}
plot(mcmc(m.jitter.19$p.beta.theta.samples))
```


```{r}
m.jitter.20 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_j_3 , "w" = 0.05)) 

#
saveRDS(m.jitter.20, "m_jitter_20.rds")
```


```{r}
names(m.jitter.20)
dim(m.jitter.20$p.w.samples) #1150 #a number for each location...
summary(as.numeric(m.jitter.20$p.w.samples))

dim(m.jitter.20$p.beta.theta.samples)

dim(m.jitter.20$p.w.knots.samples)

length(Sable_coord_binary_jitter_gdf$data)

plot(mcmc(m.jitter.20$p.beta.theta.samples))

Sable_coord_binary_jitter_gdf$data

m.jitter.21 <- spGLM(data~Sable_geo_df_binary.LONGITUDE + Sable_geo_df_binary.LATITUDE + Sable_geo_df_binary.LONGITUDE:Sable_geo_df_binary.LATITUDE, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 5000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=beta.starting_j_3, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0,0,0,0),c(1000,1000,1000,1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= beta.tuning_j_3 , "w" = 0.05)) 

#
saveRDS(m.jitter.21, "m_jitter_21.rds")
plot(mcmc(m.jitter.21$p.beta.theta.samples))

```


w's supposed to be on a grid across the region, depending on how few, how many
->

Suggestion:
make a new dataset with a centered lat... (probably not the problem)
Make a new dataset with centered longitdues, -1 to pos 1. Lat will go from -1 to pos 1.
--> center lat. Divide so it ranges from -1 to pos 1 or -2 to pos 2
Make new lat variable = (lat#-mean of all lat)/sd of lat. DO this for long as well!!!! DO THIS!!
--> new numbers for the tuning proposals for these two betas.
--> maybe I shouldnt use the same tuning proposals for lat and long
---> try to get relative size of tuning proposals to be the same. This is why we're trying this
---> DONT DO THiS, nevermind

Question:
Is my WLS exponsential varigram a good-enough fit?
How to include



MOVING FOWARD:
PLAN A: Bug Andrew O. Finley inleya@msu.edu
PLAN B: Hurdle or zero-inflated model spatial model
PLAN C: 



RESULTS!! #what to do next to get a smoothed surface? See hw 6
Use jitter 21
how to plot an spGLM result??
use spBayes funtion or Oneplot?
```{r}
X <- as.matrix(cbind(4, rnorm(1150))) #needs to be 4 by 1150?
mod.21.pred <- spPredict(m.jitter.21, pred.coords = my_grid, pred.covars = m.jitter.21$p.beta.theta.samples)
class(m.jitter.21)
```


Load one plot:
```{r}

```


RESULTS with 21
```{r}
#use Sp predict to predict coords?
burnin <- 500
#thin=10
X.ho <- matrix(data=1, nrow=1150, ncol=4)

pred <- spPredict(m.jitter.21, pred.coords = m.jitter.21$coords, pred.covars = matrix(nrow=1150, ncol=4), start = burnin, verbose = T)

pred2 <- spPredict(m.jitter.21, pred.coords = m.jitter.21$coords, pred.covars = matrix(data=1,nrow=1150, ncol=4), start = burnin, verbose = T, thin=10)
class(pred2)
names(pred2)
pred2$p.w.predictive.samples #holy shit it worked!!!!

dim(m.jitter.21$p.beta.theta.samples)
#shit, looks like this is going to take a while to get my predictions
#save that prediction
#write.csv(pred2, "predicted_sable_logistic2.csv")

pred_bind <- cbind(
  t(pred$p.w.predictive.samples),
  t(pred$p.y.predictive.samples)
) 


#example: 
library(tidybayes)
library(dplyr)
tidybayes::gather_samples(pred_bind, lambda[i]) %>%
  full_join(
    mutate(polio, i = 1:n()),
    by = "i"
  ) %>%
    ungroup() %>%
    ggplot(aes(x=year, y=estimate)) +
      tidybayes::stat_lineribbon(alpha=0.5) +
      geom_point(data=polio, aes(y=cases)) +
      facet_wrap(~term, scales = "free_y")

#try one_plot

```

Plot things
```{r}
library(MBA)
?mba.surf #a smoother
#use mba.suft, and image or image.plot
#recover samples?

############
#example code: from the more recent spbayes paper
#w.hat <- apply(m.i$p.w.recover.samples, 1, median)
#w.hat.surf <- mba.surf(cbind(coords, w.hat), no.X = res, no.Y = res, extend = TRUE)$xyz.est
#par(mar = c(5,5,0.2,0.2), cex.lab = 2, cex.axis = 2)
#image.plot(w.hat.surf, xlab = "Easting", ylab = "Northing", xaxs = "r", yaxs = "r", col = col)


#plot the predicted
names(pred)
library(ggplot2)
library(viridis)
ggplot() + aes(color=pred$p.w.predictive.samples, x=Longitude, y=latitude)+
  geom_point() + scale_color_viridis()
#the predicted graph

#teh original graph
ggplot() + aes(color=data)


```

Sample code attempt: crap
```{r}
x <- as.matrix(rep(1,1150),)
weights <- rep(1, 1150)
coords <- matrix(Sable_coord_binary_jitter_gdf$coords, ncol=2)
#weights[coords[,1]>mean(coords[,1])] <- 10

burn.in <- 0.9*1150
sub.samps <- burn.in:1150

print(summary(window(m.jitter.21$p.beta.theta.samples, start=burn.in)))

beta.hat <- m.jitter.21$p.beta.theta.samples[sub.samps,"(Intercept)"]
w.hat <- m.jitter.21$p.w.samples[,sub.samps]

p.hat <- 1/(1+exp(-(x%*%beta.hat+w.hat))) 

y.hat <- apply(p.hat, 2, function(x){rbinom(n=1150, prob=p.hat, size=1)}) #altered maybe in a bad way

y.hat.mu <- apply(y.hat, 1, mean)
y.hat.var <- apply(y.hat, 1, var)

##Take a look
par(mfrow=c(1,2))
surf <- mba.surf(cbind(coords,y.hat.mu),no.X=100, no.Y=100, extend=TRUE)$xyz.est
image(surf, main="Interpolated mean of posterior rate\n(observed rate)")
contour(surf, add=TRUE)
text(coords, label=paste("(",y,")",sep=""))

surf <- mba.surf(cbind(coords,y.hat.var),no.X=100, no.Y=100, extend=TRUE)$xyz.est
image(surf, main="Interpolated variance of posterior rate\n(observed #
of trials)")
contour(surf, add=TRUE)
text(coords, label=paste("(",weights,")",sep=""))
```


Ok ok. so pred2 is gold. Pred2 worked.
Let's make pred 3 with NO thinning next....I'll run this later
```{r}
pred3 <- spPredict(m.jitter.21, pred.coords = m.jitter.21$coords, pred.covars = matrix(data=1,nrow=1150, ncol=4), start = burnin, verbose = T)
```

For spGLM and spMisalignGLM the p.y.predictive.samples matrix holds posterior predictive samples 1/1+(exp(-x(s)'B-w(s))) and exp(x(s)'B+w(s)) for family binomial and poisson, respectively. Here s indexes the prediction location, B is the vector of regression coefficients, and w is the associated spatial random spatial effect. These values can be fed directly into rbinom or rpois to generate the realization from the respective distribution.

Ok so...maybe the rbinom will work now on the p.y.preditive samples...
```{r}
pred2$p.w.predictive.samples
pred2$p.y.predictive.samples

y.hat <- apply(pred2$p.y.predictive.samples, 1 , mean)
quant <- (function(x){quantile(x, prob=c(0.025, 0.5, 0.975))})
#y.hat <- apply(pred2$p.y.predictive.samples, 1, quant)

#plot(y, y.hat[2,], pch=19, cex=0.5, xlab="observed y", ylab="predicted y")
#arrows(y[-mod], y.hat[2,-mod], y[-mod], y.hat[1,-mod], angle=90, length=0.05)
#arrows(y[-mod], y.hat[2,-mod], y[-mod], y.hat[3,-mod], angle=90, length=0.05)

ggplot() + aes(x=Sable_coord_binary_jitter_gdf$data, y=y.hat) + geom_point()

df <- data.frame(Sable_coord_binary_jitter_gdf)
df$data <- as.factor(df$data)

y.vec<- as.vector(y.hat)
y.hat.factor <- as.factor(y.vec)
length(y.vec)
y.vector.real <- rep(1,1150)
y.hat.fac<-as.factor(y.vector.real)



ggplot(df) +  aes(x=Sable_geo_df_binary.LONGITUDE, y=Sable_geo_df_binary.LATITUDE, color=y.hat.fac) +geom_point() + scale_color_viridis_d() +
  labs(x="Centered Longitude", y="Latitude") +
  theme_bw()

ggplot(df) +  aes(x=Sable_geo_df_binary.LONGITUDE, y=Sable_geo_df_binary.LATITUDE, color=data) +geom_point() + scale_color_viridis_d(limits=c("1", "0")) +
  labs(x="Centered Longitude", y="Latitude") +
  theme_bw()

range(y.hat)
```
Let's run a prediction with the fixed identity matrix
```{r}
Iden <- matrix(data=1, nrow=1150, ncol=4)
Iden[,2] <- Sable_coord_binary_jitter_gdf$coords[,1]
Iden[,3] <- Sable_coord_binary_jitter_gdf$coords[,2]
Iden[,4] <- Sable_coord_binary_jitter_gdf$coords[,1]*Sable_coord_binary_jitter_gdf$coords[,2]


pred3 <- spPredict(m.jitter.21, pred.coords = m.jitter.21$coords, pred.covars = Iden, start = burnin, verbose = T, thin=10)

pred3$p.y.predictive.samples

pred4 <- spPredict(m.jitter.21, pred.coords = m.jitter.21$coords, pred.covars = Iden, start = burnin, verbose = T)

#write.csv(pred3, "pred3.csv") #the successful one, thinned
#write.csv(pred4, "pred4.csv") #the successful one

pred4$p.y.predictive.samples #ok that failed. But pred3 is ok... might try to run again


#use pred3 for now. That's the one that worked.
#what to do with pred3 to produce a map

#can be fed into rbinom to create a realization of the distribution
##or try mba surf or whatever
set.seed(1)
results_3 <- rbinom(1150, size=1, prob=pred4$p.y.predictive.samples)
results_3_fac <- as.factor(results_3)
ggplot(df) +  aes(x=Sable_geo_df_binary.LONGITUDE, y=Sable_geo_df_binary.LATITUDE, color=results_3_fac) +geom_point() + scale_color_viridis_d(limits=c("1", "0"), name="data") +
  labs(x="Centered Longitude", y="Latitude") +
  theme_bw()
```
MBA surf
going to work on other things. Here is what is important:
df(right now for the ggplot)
Sable_coord_binary_jitter_gdf
pred3
To do next:
MBA surf. Other stuff too?
```{r}
write.csv(Sable_coord_binary_jitter_gdf, "Sable_coord_binary_jitter_gdf.csv")
```



Surfs up! smoothed map time!!
```{r}
library(MBA)
?mba.surf
#might need non-factor resutls, but trying it first with factor results
surfs_up_pred <- mba.surf(cbind(Sable_coord_binary_jitter_gdf$coords,results_3_fac),
                          no.X=100,
                          no.Y=100,
                          extend=T
                          )
est_surf_pred <- surfs_up_pred$xyz.est
image(est_surf_pred)
points(Sable_coord_binary_jitter_gdf$coords)
contour(est_surf_pred, add=T ) #this is it!! But might try without the points

image(est_surf_pred)
contour(est_surf_pred, add=T ) #this is the predticed without he points

#let's look at og data surf with the points

surfs_up_obs <- mba.surf(cbind(Sable_coord_binary_jitter_gdf$coords,df$data),
                          no.X=100,
                          no.Y=100,
                          extend=T
                          )
est_surf_obs <- surfs_up_obs$xyz.est

image(est_surf_obs)
points(Sable_coord_binary_jitter_gdf$coords)
contour(est_surf_obs, add=T )


#let's look at an og data surf without the points
image(est_surf_obs)
contour(est_surf_obs, add=T )  


#surfs
```



Can I add alaska background lines/shape to my map?
```{r}

```

Effective sample size
```{r}
library(coda)
effectiveSize(m.jitter.21$p.beta.theta.samples)
#that's really bad. Let's not include that....
cor(m.jitter.21$p.beta.theta.samples)
```

04/25/22
Try with no trend: just intercept?
Same WLS, need new GLM
```{r}
glm_int <- glm(data ~ 1, data=Sable_coord_binary_jitter_gdf, family=binomial(link="logit"))
summary(glm_int)
names(glm_int)
glm_int$coef
int_glm <- glm_int$coef

#need that weird transpose from the example
#beta.tun.int <- 

m.jitter.int <- spGLM(data~1, family="binomial", coords=Sable_coord_binary_jitter_gdf$coords, data=Sable_coord_binary_jitter_gdf, n.samples = 1000, cov.model="exponential", #change to 5000 eventually
             starting=list("phi"=sable_WLS_phi_j,"sigma.sq"=sable_WLS_sigsq_j, "beta"=int_glm, 
                           "w"= 0), #using 0, like in the example, because this w is NOT omega
             priors=list("phi.Unif"=c(0.01, 5), "sigma.sq.IG"=c(2,1),                                                                                                     "beta.Normal"=list(c(0),c(1000))), #gonna want to make these 1000...
             tuning=list("phi"=0.05, "sigma.sq"=0.05, "beta"= 0.05, "w" = 0.05)) 


library(coda)
#plot
```

Traceplots
```{r}
#plot them traces for m.jitter.21
plot(mcmc(cbind(m.jitter.21$p.beta.theta.samples)), density=F)
#this will be figure 3

```

